---
title: "Supplementary Appendix 3"
author: "NCPE"
geometry: margin=2cm
output: 
  bookdown::pdf_document2:
      extra_dependencies: ["float","longtable","booktabs"]
      number_sections: yes
      
---

```{r setup, include=FALSE}

library(tidyverse)
library(flexsurv)
library(GGally)
library(ggpubr)
library(kableExtra)
library(formatR)

knitr::opts_chunk$set(echo = FALSE,
                      message= FALSE, 
                      warning=FALSE, 
                      cache=TRUE,
                      tidy=TRUE,
                      dev = 'png',
                      tidy.opts=list(width.cutoff=60),
                      fig.pos = "H", 
                      out.extra = "",
                      latex_options = "HOLD_position"
                      )

set.seed(12345)

```


```{r run}
# Fit the models etc
source("setup.R")
source("run_all.R")
```



# Importance sampling: model diagnostics, parameter distributions, and survival estimates

The tables and figures in this section show survival curve parameter estimates, covariance structures, importance sampling diagnostics, and (probabilistic) survival curve predictions, obtained from trial data only estimated using maximum likelihood estimation (MLE), and combining trial data with external information using the importance sampling algorithm described in the methods section of the main article and in Appendix 1.

```{r captions}

captions1<-c(paste0("Effective sample size (ESS), corresponding to 5000 (weighted) parameter samples, plotted as a function of the tempering parameter alpha, for the ",distributions," distribution. The vertical line shows the optimal value of alpha identified by the algorithm."))
captions2<-c(paste0("Survival curve parameter distributions and covariance estimates obtained using trial data only (MLE) and combining trial data with external information  (IS), ",distributions," distribution."))
captions3<-c(paste0("Survival curves (median and 95% confidence intervals) obtained using trial data only (MLE) and combining trial data with external information (IS), ",distributions," distribution. Vertical bar indicates 95% probability interval for 5 year OS based on external information."))

captions<-character(0)

for (i in 0:5){
  
  j<-1+3*i
  
  captions[j]<-captions1[i+1]
  captions[j+1]<-captions2[i+1]
  captions[j+2]<-captions3[i+1]
  
}
## remove if this makes things worse AW 
captions <- matrix(captions, nrow = 3, ncol = 6)
captions <- t(captions)
captions <- as.numeric(captions)
```


```{r plotsonly, results='hide', error=FALSE, message=FALSE, warning=FALSE,cache=T}
invisible(
for (dist in dists)
{
if(!file.exists(paste0("./output/plots_",dist))){
  plots<-is_surv_viz_gg(is.models[[dist]],tseq2,tstar,what=1:3,dist=dist)
  saveRDS(plots,file=paste0("./output/plots_",dist))} else
{
    plots<-readRDS(paste0("./output/plots_",dist))
  }

}
)
```

## Comparisons of parameter estimates with and without external information 

The below tables compare parameter estimates and covariance matrices obtained from trial data only with those obtained combining trial data with external information, for each of the survival models under consideration. 


```{r paramcomp, results='asis',fig.cap=captions, dpi=300, fig.height=4, fig.width=6}
names(is.models[["exponential"]][["orig"]]$coefficients) <- names(is.models[["exponential"]]$post_mean) <- "rate"

dimnames(is.models[["exponential"]][["orig"]]$cov) <- dimnames(is.models[["exponential"]]$post_cov) <- list("rate", "rate")

for (dist in dists)
{

    print(knitr::kable(
      list(is.models[[dist]][["orig"]]$coefficients,is.models[[dist]]$post_mean),
      caption=paste0("Parameter estimates obtained using trial data only (left) and combining trial data with external information (right): ",distributions[dist]," distribution.")))

  
  cat('\n\n') 
  
  print(knitr::kable(
    list(is.models[[dist]][["orig"]]$cov, is.models[[dist]]$post_cov),
    caption=paste0("Covariance matrices obtained using trial data only (left) and combining trial data with external information (right): ",distributions[dist], " distribution.")))
  
cat('\n\n') 
}
  cat('\\newpage')
  
```


## Effective sample size plots  

The below plots show the estimated effective sample size (ESS), corresponding to 5000 (weighted) parameter samples, plotted as a function of the tempering parameter alpha, for the survival distributions under consideration. The vertical line shows the optimal value of alpha identified by the algorithm. Larger ESS values are better. 


```{r ESSplots, results='asis',fig.cap=captions, dpi=300, fig.height=4, fig.width=6}

for (dist in dists)
{

  plots<-readRDS(paste0("./output/plots_",dist))
  print(plots[[1]]) 
  # +labs(title=NULL)) ## I think including the titles is better?? 

  cat('\n\n')
}


```

## Distribution comparison plots 

The below plots show the estimated effective sample size (ESS), corresponding to 5000 (weighted) parameter samples, plotted as a function of the tempering parameter alpha, for the survival distributions under consideration. The vertical line shows the optimal value of alpha identified by the algorithm. Larger ESS values are better. 

```{r distplots, results='asis',fig.cap=captions, dpi=300, fig.height=4, fig.width=6}
  
for (dist in dists) 
{

  plots<-readRDS(paste0("./output/plots_",dist))
  print(plots[[2]])
  # +labs(title=NULL))

  cat('\n\n')
}

 

```

## Survival comparison plots 

The below plots show the estimated survival curves (median and 95% confidence intervals) obtained using trial data only and combining trial data with external information, for the survival distributions under consideration. Vertical bars indicate the 95% probability interval for 5 year OS based on external information.

```{r survplots, results='asis',fig.cap=captions, dpi=300, fig.height=4, fig.width=6}
  
for (dist in dists)
{

  plots<-readRDS(paste0("./output/plots_",dist))
  print(plots[[3]])
  # +labs(title=NULL))

  cat('\\newpage')
}

 

```

<!-- ```{r allplots, results='asis',fig.cap=captions, dpi=300, fig.height=4, fig.width=6} -->

<!-- for (dist in dists[-1]) -->
<!-- { -->

<!--     print(knitr::kable( -->
<!--       list(is.models[[dist]][["orig"]]$coefficients,is.models[[dist]]$post_mean), -->
<!--       caption=paste0("Parameter estimates obtained using trial data only (left) and combining trial data with external information (right): ",distributions[dist]," distribution."))) -->


<!--   cat('\n\n')  -->



<!--   print(knitr::kable( -->
<!--     list(is.models[[dist]]$post_cov,is.models[[dist]][["orig"]]$cov), -->
<!--     caption=paste0("Covariance matrices obtained using trial data only (left) and combining trial data with external information (right): ",distributions[dist], " distribution."))) -->

<!-- cat('\n\n')  -->


<!--   plots<-readRDS(paste0("./output/plots_",dist)) -->



<!--   print(plots[[1]]+labs(title=NULL)) -->

<!--   cat('\n\n')  -->

<!--   print(plots[[2]]+labs(title=NULL)) -->

<!--   cat('\n\n')  -->

<!--   print(plots[[3]]+labs(title=NULL)) -->

<!--   cat('\\newpage')  -->
<!-- } -->



<!-- ``` -->



```{r, cache=F}
#knitr::knit_exit()
```

## Parameter variability

Table \@ref(tab:parameffects) shows 'Generalised variance,' i.e., determinants of variance-covariance matrices, of the parameter estimates obtained using trial data only (MLE) and combining trial data withe external information (IS). Formally, the generalised variance gives the area of the 95% highest density ellipse and can be interpreted as a 1-parameter measure of parameter uncertainty. 

<!-- , see e.g., https://stats.stackexchange.com/questions/12762/measure-of-spread-of-a-multivariate-normal-distribution . Not sure about citing stack exchange in an appendix to a paper??  AW -->
<!-- Now that I think about it, I guess that the determinant is just the trace of the covariance (i.e., sum of the variances), offset by the covariance entries on the off diagnoal? Is my intuition ok here?? AW  -->



```{r parameffects,results='asis'}

param.uncertainty<-data.frame("Distribution"=dists,"Distance"=rep(NA,6),"MLE Variance"=rep(NA,6),"IS Variance"=rep(NA,6),"Var. Ratio"=rep(NA,6))

for (i in seq_along(dists))
{
  param.uncertainty[i,2]<-sqrt(sum((is.models[[dists[i]]][["orig"]]$coefficients-is.models[[dists[i]]]$post_mean)^2))
  param.uncertainty[i,3]<-det(as.matrix(is.models[[dists[i]]][["orig"]]$cov))
  param.uncertainty[i,4]<-det(as.matrix(is.models[[dists[i]]]$post_cov))
  param.uncertainty[i,5]<-param.uncertainty[i,4]/param.uncertainty[i,3]
  
}

cat('\n\n')

knitr::kable((param.uncertainty),
             caption="Generalised variance of model parameter distributions. Comparison between parameter estimates obtained using trial data only (MLE) and combining trial data with external information (IS).",
             booktabs=TRUE) %>%
             kableExtra::kable_styling(latex_options = "hold_position")
             

```


# Comparison with expertsurv output

```{r morecaptions}
captions4<-paste0("Survival curves (median and 95 percent confidence intervals) obtained by combining trial data with external information using the importance sampling (IS) and expertsurv methods, ",distributions," distribution")

captions5<-c(paste0("Survival curve parameter distributions and covariance estimates obtainedby combining trial data with external information using the importance sampling (IS) and expertsurv methods, ",distributions," distribution"))

```


## Plots of survival curves over time

Figures \@ref(fig:expertsurvIS-1) to \@ref(fig:expertsurvIS-6) show comparisons of survival curves (median and 95% confidence/credible intervals) obtained from the importance sampling method described in the paper, with those obtained from the fully Bayesian approach of Cooney & White implemented in the expertsurv package.

Survival curves for the exponential, Weibull, Gompertz and log-logistic models were almost identical between the two methods. For the log-normal and generalised gamma models (Figures \@ref(fig:expertsurvIS-4) and \@ref(fig:expertsurvIS-6)), there was a small but notable shift in the parameter distribution and corresponding survival estimates. The reason for this is not immediately clear. However, in both cases the impact was small and mostly evident in the tails of the survival distributions.

```{r expertsurvIS, results='asis',fig.cap=captions4, dpi=300}
source("comp_expertsurv.R")
# By curve type
for(i in 1:6)
{
  g<-ggplot(data=surv.list[[i]],aes(x=time,colour=Method,fill=Method))+
    geom_line(aes(y=S_median),lwd=1)+
    geom_ribbon(aes(ymin=S_lower,ymax=S_upper),alpha=0.1,linetype="dashed")+
    labs(y="S(t)",x="time (t)",title=paste0("Survival - ",distributions[dists[i]]))
  
  cat('\n\n') 
  print(g)
  cat('\n\n') 
  
}
```

### Comparisons of parameter distributions

Figures \@ref(fig:plotparams-1) to \@ref(fig:plotparams-6) compare survival curve parameter distributions obtained from the importance sampling method described in the paper, with those obtained from the fully Bayesian approach of Cooney & Whilte implemented in the expertsurv package.  Again, some differences are evidence between the two methods for the log-normal and generalised gamma models (Figures \@ref(fig:plotparams-4) and \@ref(fig:plotparams-6)).

```{r plotparams,message=F, warning=F, fig.cap=captions5, dpi=300}

library(GGally)
cat('\n\n') 

for(i in dists){
# Extract parameter draws from the two types of model fit and merge
  df1<-data.frame(is.sims[[i]][["sims.mvn"]],"Method"=rep("IS",5000))
  
  df2<-data.frame(exs.sims[[i]][["sims.mvn"]],"Method"=rep("expertsurv",5000))
  
  names(df2)<-names(df1)
  
  df<-bind_rows(df1,df2)
  
  cat('\n\n') 
  
print(ggpairs(df, aes(colour = Method, alpha = 0.4),
              columns = 1:(ncol(df)-1),
              title=paste0("Parameters - ",distributions[i])))
cat('\n\n') 
}
  
``` 

## Comparisons of AUC distributions

Table \@ref(tab:AUC) and Figure \@ref(fig:AUC) compare probabilistic estimates of area under the curve between the two methods. In general, AUC measures (population) mean survival extrapolated over a lifetime, though in these examples survival is capped at 100 years since some curves plateau.

```{r AUC,  results='asis', fig.cap="Density plots comparing probabilistic total area under the curve (mean OS) distributions obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, with those obtained from the fully Bayesian method of Cooney and White implemented in the expertsurv package. "}

cat('\n\n') 

#AUC for the models

auc.is<-data.frame(matrix(NA,nrow=5000,ncol=6))
names(auc.is)<-dists
auc.exs<-auc.is


for(dist in dists)
{
  auc.is[dist]<-is.sims[[dist]][["AUC"]]
  auc.exs[dist]<-exs.sims[[dist]][["AUC"]]
}

auc.is["Method"]="IS"
auc.exs["Method"]="expertsurv"

auc.df2<-bind_rows(auc.is,auc.exs) %>%
  pivot_longer(1:6,
               names_to="Distribution",
               values_to="AUC")

auc.summary2<-auc.df2 %>%
  group_by(Distribution,Method) %>%
  summarise(mean=mean(AUC),
            sd=sd(AUC),
            median=median(AUC),
            lwr.95=quantile(AUC,0.025),
            upr.95=quantile(AUC,0.975)) 

# Same for AUC

auc.print2 <- auc.summary2 %>%
  pivot_wider(names_from = Method,values_from = 3:7) %>% 
  mutate(mean_diff=mean_IS-mean_expertsurv,
         var.ratio=sd_IS^2/sd_expertsurv^2,
         Distribution=str_replace_all(Distribution,distributions)) %>% 
  mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)))  %>%
  mutate(expertsurv=paste0(mean_expertsurv," (",lwr.95_expertsurv,", ",upr.95_expertsurv,")"),
         IS=paste0(mean_IS," (",lwr.95_IS,", ",upr.95_IS,")"),
         Var.Ratio=var.ratio
  ) %>%
  ungroup() %>%
  select(Distribution,expertsurv,IS,Mean.Diff=mean_diff,Var.Ratio) 

knitr::kable(auc.print2,
             caption = "Comparison of probabilistic total area under the curve (mean lifetime OS) estimates, obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, and with the fully Bayesian method of Cooney and White implemented in the expertsurv package. Values shown are probabilstic mean and 95 percent confidence/credible intervals.")

# AUC density plot
# No truncation of AUC values but truncation of graph
library(ggridges)

cat('\n\n') 

g.auc<-ggplot(data=auc.df2,aes(x=AUC,y=Distribution,colour=Method,fill=Method))+
  geom_density_ridges(alpha=0.5,scale=0.95)+
  scale_fill_manual(values=cbPalette[2:3])+
  scale_colour_manual(values=cbPalette[2:3])+
  labs(x="AUC (Mean lifetime OS, Months)",y="Density")+
  xlim(0,360)+
  theme(legend.position = "bottom")

g.auc

cat('\n\n') 

```




## Comparisons of 5-year OS

Table \@ref(tab:ststar) and Figure \@ref(fig:ststar) compare probabilistic estimates of 5-year OS obtained from the two methods.


```{r ststar,  results='asis', fig.cap = "Density plots comparing 5-year OS estimates, obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, with those obtained from the fully Bayesian method of Cooney and White implemented in the expertsurv package."}
library(scales)

surv.tstar2<-data.frame()

for (dist in dists)
{
  tmp.df<-data.frame("expertsurv"=unname(exs.sims[[dist]][["s.tstar"]]),
                     "IS"=unname(is.sims[[dist]][["s.tstar"]]),
                     "dist"=dist)
  surv.tstar2<-bind_rows(surv.tstar2,tmp.df)
  rm(tmp.df)
}

surv.tstar.long2<-pivot_longer(surv.tstar2,cols=1:2,values_to = "S.tstar") %>%
  mutate(Output=factor(name,levels=c("IS","expertsurv")),
         Distribution=str_replace_all(dist,distributions))

surv.tstar.summary2<-surv.tstar.long2 %>% 
  group_by(Distribution,Output) %>%
  summarise(mean=mean(S.tstar),
            median=median(S.tstar),
            sd=sd(S.tstar),
            lwr.95=quantile(S.tstar,0.025),
            upr.95=quantile(S.tstar,0.975)) %>%
  ungroup %>%
  mutate(across(where(is.numeric),~percent(.x,accuracy=0.01))) 
cat('\n\n') 

knitr::kable(surv.tstar.summary2,
             caption = "Comparison of 5-year OS estimates, obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, with the fully Bayesian method of Cooney and White implemented in the expertsurv package.")

g.ststar2<-ggplot(data=surv.tstar.long2,aes(x=S.tstar,y=Distribution,colour=Output,fill=Output))+
  geom_density_ridges(alpha=0.5,scale=1)+
  scale_fill_manual(values=cbPalette)+
  scale_colour_manual(values=cbPalette)+
  labs(x="5-year landmark OS",y="Density")

cat('\n\n') 
g.ststar2
cat('\n\n') 
```
