---
title: "Supplementary Appendix 3"
geometry: margin=2cm
output: 
  bookdown::pdf_document2:
      extra_dependencies: ["float","longtable","booktabs"]
      number_sections: yes
      fig_caption: true
      
---

```{r setup, include=FALSE}

library(tidyverse)
library(flexsurv)
library(GGally)
library(ggpubr)
library(kableExtra)
library(formatR)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE,
                      message= FALSE, 
                      warning=FALSE, 
                      cache=TRUE,
                      tidy=TRUE,
                      dev = 'png',
                      tidy.opts=list(width.cutoff=60),
                      fig.pos = "H", 
                      out.extra = "",
                      latex_options = "HOLD_position"
                      )

set.seed(12345)

```


```{r run}
# Fit the models etc
source("setup.R")
source("run_all.R")
```


```{r captions}

captions1<-c(paste0("Effective sample size (ESS), corresponding to 5000 (weighted) parameter samples, plotted as a function of the tempering parameter alpha, for the ",distributions," distribution. The vertical line shows the optimal value of alpha identified by the algorithm."))
captions2<-c(paste0("Survival curve parameter distributions and covariance estimates obtained using trial data only and combining trial data with external information, ",distributions," distribution."))
captions3<-c(paste0("Survival curves (median and 95% confidence intervals) obtained using trial data only and combining trial data with external information, ",distributions," distribution. Vertical bar indicates 95% probability interval for 5 year OS based on external information."))

captions<-character(0)

for (i in 0:5){
  
  j<-1+3*i
  
  captions[j]<-captions1[i+1]
  captions[j+1]<-captions2[i+1]
  captions[j+2]<-captions3[i+1]
  
}
## remove if this makes things worse AW 
captions <- matrix(captions, nrow = 3, ncol = 6)
captions <- t(captions)
captions <- as.numeric(captions)
```


```{r plotsonly, results='hide', error=FALSE, message=FALSE, warning=FALSE,cache=T}
# Create the plots if needed, otherwise load them
invisible(
for (dist in dists)
{
if(!file.exists(paste0("./output/plots_",dist))){
  plots<-is_surv_viz_gg(is.models[[dist]],tseq2,tstar,what=1:3,dist=dist)
  saveRDS(plots,file=paste0("./output/plots_",dist))} else
{
    plots<-readRDS(paste0("./output/plots_",dist))
  }

}
)
```

# Plots of survival curves



The below plots show the estimated survival curves (pointwise probabilistic median and 95% confidence intervals) obtained using trial data only and combining trial data with external information, for the survival distributions under consideration. Vertical bars indicate the 95% probability interval for 5 year OS based on external information.
                                                    
```{r survplots, results='asis',fig.cap=captions3, dpi=300, fig.height=3.5, fig.width=7}

for (dist in dists)
{
plots<-readRDS(paste0("./output/plots_",dist))
print(plots[[3]])
# +labs(title=NULL))
#cat('\\newpage')
}
```
                                                   



# Probabilistic distributions of 10- and 20-year OS

```{r}

## This code summarises and plots probability density of survival estimates at time 10 years and 20 years from the prior, likelihood, and 'posterior'

surv.s10s20<-data.frame()

for(dist in dists)
{
  # Extract parameter sims
  param.sims.mle<-mle.sims[[dist]][["sims.nat"]]
  param.sims.is<-is.sims[[dist]][["sims.nat"]]
  
  # Predict survival
  
  df.mle<-data.frame(sapply(c(120,240),function(x){s_fun(t=x,dist=dist,trans_pars=param.sims.mle)}))
  names(df.mle)<-c("S.10","S.20")
  df.mle[,"method"]<-"MLE"
  
  df.is<-data.frame(sapply(c(120,240),function(x){s_fun(t=x,dist=dist,trans_pars=param.sims.is)}))
  names(df.is)<-c("S.10","S.20")
  df.is[,"method"]<-"IS"
  
  df.st<-rbind.data.frame(df.mle,df.is)
  df.st[,"distribution"]<-dist
  
  # Combine into existing datafram
  
  surv.s10s20<-rbind.data.frame(surv.s10s20,df.st)
  
}

surv.s10s20.long<-pivot_longer(surv.s10s20,cols=1:2,values_to = "S.tstar") %>%
  mutate(Distribution=str_replace_all(distribution,distributions),
         Timepoint=str_replace_all(name,c("S.10"="10 years","S.20"="20 years")),
         Output=str_replace_all(method,c("MLE"="Trial data only","IS"="Trial data with external information")))


###


surv.s10s20.summary<-surv.s10s20.long %>% 
  group_by(Distribution,Timepoint,method) %>%
  summarise(mean=mean(S.tstar),
            median=median(S.tstar),
            sd=sd(S.tstar),
            lwr.95=quantile(S.tstar,0.025),
            upr.95=quantile(S.tstar,0.975),
            ) %>%
  ungroup()%>%
  pivot_wider(names_from=method,
              values_from=4:8) %>%
  mutate(mean_diff=mean_IS-mean_MLE,
         var_ratio=sd_IS^2/sd_MLE^2,
         across(where(is.numeric),~format(round(.x,2),nsmall=2)),
         CrI_IS=paste0("(",lwr.95_IS," to ",upr.95_IS,")"),
         CrI_MLE=paste0("(",lwr.95_MLE," to ",upr.95_MLE,")"),
  ) %>%
  select(Distribution,Timepoint,mean_MLE,median_MLE,CrI_MLE,mean_IS,median_IS,CrI_IS,mean_diff,var_ratio)
  
#write.csv(surv.s10s20.summary,"./output/summary.s10s20.csv")


kbl(surv.s10s20.summary,
    caption = "Summary of probabilistic distributions of 10- and 20-year OS, estimated using Trial Data Only via MLE, and combining Trial Data with External information via the Importance Sampling method",
    col.names = c("Distribution", "Timepoint", rep(c("Mean","Median","95% CrI"),2),"Mean Diff.","Var. Ratio"),
    booktabs=T) %>%
  kable_styling(latex_options = c("HOLD_position","striped")) %>%
  add_header_above(c(" "=2,
                     "Trial Only"=3,
                     "Trial + External" =3,
                     "Difference"=2))

```

```{r, fig.dim=c(7,9),fig.cap = "Density plots of probabilistic distributions of 10- and 20-year OS, estimated using Trial Data Only via MLE, and combining Trial Data with External information via the Importance Sampling method",dpi=300}

# Version in use
g.s10s20<-ggplot(data=surv.s10s20.long,aes(x=S.tstar,fill=Output,height=after_stat(scaled)))+
  geom_density(alpha=0.5,#scale=1,
                      #rel_min_height=0.01,
                       trim=TRUE,
                      #stat="scaled",
                      bounds=c(0,1)
  )+
  #geom_histogram()+
  scale_fill_manual(values=cbPalette[2:3])+
  scale_colour_manual(values=cbPalette[2:3])+
  labs(x="OS probability",y="Density")+
  facet_wrap(Distribution~Timepoint,ncol=2,scales="free_y",
             labeller = 
               label_wrap_gen(multi_line=FALSE),
             strip.position="top"
               #labeller()
             )+
 # facet_grid(Distribution~Timepoint,scales="free",shrink=TRUE)+
  theme(legend.position = "bottom",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


print(g.s10s20)


```

# Importance sampling: model diagnostics, parameter distributions, and survival estimates

The tables and figures in this section show survival curve parameter estimates, covariance structures, importance sampling diagnostics, and (probabilistic) survival curve predictions, obtained from trial data only estimated using maximum likelihood estimation (MLE), and combining trial data with external information using the importance sampling algorithm described in the methods section of the main article and in Appendix 1.



## Comparisons of parameter estimates with and without external information 

The below tables compare parameter estimates and covariance matrices obtained from trial data only with those obtained combining trial data with external information, for each of the survival models under consideration. 


```{r paramcomp, results='asis', dpi=300, fig.height=4, fig.width=6}
names(is.models[["exponential"]][["orig"]]$coefficients) <- names(is.models[["exponential"]]$post_mean) <- "rate"

dimnames(is.models[["exponential"]][["orig"]]$cov) <- dimnames(is.models[["exponential"]]$post_cov) <- list("rate", "rate")

for (dist in dists)
{

    kable(
      list(is.models[[dist]][["orig"]]$coefficients,is.models[[dist]]$post_mean),
      caption=paste0("Parameter estimates obtained using trial data only (left) and combining trial data with external information (right): ",distributions[dist]," distribution."),
      booktabs=T) %>%
    kable_styling(latex_options=c("striped","HOLD_position")) %>%
    print()

  
  
  cat('\n\n') 
  
  kable(
    list(is.models[[dist]][["orig"]]$cov, is.models[[dist]]$post_cov),
    caption=paste0("Covariance matrices obtained using trial data only (left) and combining trial data with external information (right): ",distributions[dist], " distribution."),
    booktabs=TRUE) %>%
      kable_styling(latex_options=c("striped","HOLD_position")) %>%
    print()

    
  
cat('\n\n') 
}
  #cat('\\newpage')
  
```






## Effective sample size plots  

The below plots show the estimated effective sample size (ESS), corresponding to 5000 (weighted) parameter samples, plotted as a function of the tempering parameter alpha, for the survival distributions under consideration. The vertical line shows the optimal value of alpha identified by the algorithm. Larger ESS values are better. 


```{r ESSplots, results='asis',fig.cap=captions1, dpi=300, fig.height=3.5, fig.width=6}

for (dist in dists)
{
  
  plots<-readRDS(paste0("./output/plots_",dist))
  print(plots[[1]]) 
  # +labs(title=NULL)) ## I think including the titles is better?? 
  
  #cat('\n\n')
}


```

## Distribution comparison plots 

The below plots show parameter distributions and scatter plot estimates obtained using trial data only and by combining trial data with external information for the survival distributions under consideration.

```{r distplots, results='asis',fig.cap=captions2, dpi=300, fig.height=4, fig.width=7}

for (dist in dists) 
{
  
  plots<-readRDS(paste0("./output/plots_",dist))
  print(plots[[2]])
  # +labs(title=NULL))
  
  cat('\n\n')
}


```



## Parameter variability
Table \@ref(tab:parameffects) shows 'Generalised variance,' i.e., determinants of variance-covariance matrices, of the parameter estimates obtained using trial data only (MLE) and combining trial data withe external information (IS). Formally, the generalised variance gives the area of the 95% highest density ellipse and can be interpreted as a 1-parameter measure of parameter uncertainty. 
                                                    
<!-- , see e.g., https://stats.stackexchange.com/questions/12762/measure-of-spread-of-a-multivariate-normal-distribution . Not sure about citing stack exchange in an appendix to a paper??  AW -->
<!-- Now that I think about it, I guess that the determinant is just the trace of the covariance (i.e., sum of the variances), offset by the covariance entries on the off diagnoal? Is my intuition ok here?? AW  -->

```{r parameffects,results='asis'}                                                    
param.uncertainty<-data.frame("Distribution"=dists,"Distance"=rep(NA,6),"MLE Variance"=rep(NA,6),"IS Variance"=rep(NA,6),"Var. Ratio"=rep(NA,6))
                  
for (i in seq_along(dists))
{                                                      param.uncertainty[i,2]<-sqrt(sum((is.models[[dists[i]]][["orig"]]$coefficients-is.models[[dists[i]]]$post_mean)^2))
                                                      param.uncertainty[i,3]<-det(as.matrix(is.models[[dists[i]]][["orig"]]$cov))
                                                      param.uncertainty[i,4]<-det(as.matrix(is.models[[dists[i]]]$post_cov))
                                                      param.uncertainty[i,5]<-param.uncertainty[i,4]/param.uncertainty[i,3]
                                                      
}

cat('\n\n')
                                                    
knitr::kable((param.uncertainty),
caption="Generalised variance of model parameter distributions. Comparison between parameter estimates obtained using trial data only (MLE) and combining trial data with external information (IS).",
booktabs=TRUE) %>%
kableExtra::kable_styling(latex_options = c("HOLD_position","striped"))

```


# Comparison with expertsurv output

In this section we compare the outputs obtained from the importance sampling method described in the paper to that obtained using the Cooney & White method implemented in the expertsurv package. For the gen. Gamma model estimated using the expertsurv package, suitable Monte Carlo performance could not be obtained and hence results are not reported. 

```{r morecaptions}
captions4<-paste0("Survival curves (median and 95 percent confidence intervals) obtained by combining trial data with external information using the importance sampling (IS) and expertsurv methods, ",distributions," distribution")

captions5<-c(paste0("Survival curve parameter distributions and covariance estimates obtainedby combining trial data with external information using the importance sampling (IS) and expertsurv methods, ",distributions," distribution"))

# delete gen gamma

captions4<-captions4[1:5]
captions5<-captions5[1:5]

```


## Plots of survival curves over time

Figures \@ref(fig:expertsurvIS-1) to \@ref(fig:expertsurvIS-5) show comparisons of survival curves (median and 95% confidence/credible intervals) obtained from the importance sampling method described in the paper, with those obtained from the fully Bayesian approach of Cooney & White implemented in the expertsurv package.

Survival curves for the exponential, Weibull, Gompertz and log-logistic models were almost identical between the two methods. For the log-normal model (Figure \@ref(fig:expertsurvIS-4)) there was a small but notable shift in the parameter distribution and corresponding survival estimate. The reason for this is not immediately clear. However, in both cases the impact was small and mostly evident in the tail of the survival distribution.

```{r expertsurvIS, results='asis',fig.cap=captions4, dpi=300, fig.height=4, fig.width=7}
source("comp_expertsurv.R")
# By curve type
for(i in 1:5)
{
                                                      g<-ggplot(data=surv.list[[i]],aes(x=time,colour=Method,fill=Method))+
geom_line(aes(y=S_median),lwd=1)+ geom_ribbon(aes(ymin=S_lower,ymax=S_upper),alpha=0.1,linetype="dashed")+
labs(y="S(t)",x="time (t)",title=paste0("Survival - ",distributions[dists[i]]))
cat('\n\n')
print(g)
cat('\n\n')
}
```

## Comparisons of parameter distributions

Figures \@ref(fig:plotparams-1) to \@ref(fig:plotparams-5) compare survival curve parameter distributions obtained from the importance sampling method described in the paper, with those obtained from the fully Bayesian approach of Cooney & Whilte implemented in the expertsurv package.  Again, some differences are evident between the two methods for the log-normal model (Figure \@ref(fig:plotparams-4)).

```{r plotparams,message=F, warning=F, fig.height=4,fig.width=7,fig.cap=captions5, dpi=300}

library(GGally)
cat('\n\n')

for(i in dists[-6]){
# Extract parameter draws from the two types of model fit and merge
                                                      df1<-data.frame(is.sims[[i]][["sims.mvn"]],"Method"=rep("IS",5000))
df2<-data.frame(exs.sims[[i]][["sims.mvn"]],"Method"=rep("expertsurv",5000))
names(df2)<-names(df1)
df<-bind_rows(df1,df2)

cat('\n\n')
if(i == "exponential"){
      
      print(ggplot(df, aes(x = rate, colour = Method, fill = Method)) + 
        geom_density(alpha=0.4 ) +
        labs(title="Parameters - exponential"))
    } else{

print(ggpairs(df, aes(colour = Method, alpha = 0.4),
columns = 1:(ncol(df)-1), title=paste0("Parameters - ",distributions[i])))
    }
cat('\n\n')
}
```

## Comparisons of AUC distributions

Table \@ref(tab:AUC) and Figure \@ref(fig:AUC) compare probabilistic estimates of area under the curve between the two methods. In general, AUC measures (population) mean survival extrapolated over a lifetime, though in these examples survival is capped at 100 years since some curves plateau.

```{r AUC,  results='asis', fig.cap="Density plots comparing probabilistic total area under the curve (mean OS) distributions obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, with those obtained from the fully Bayesian method of Cooney and White implemented in the expertsurv package. "}

cat('\n\n')

#AUC for the models
auc.is<-data.frame(matrix(NA,nrow=5000,ncol=6))
names(auc.is)<-dists
auc.exs<-auc.is

for(dist in dists[-6])
{
auc.is[dist]<-is.sims[[dist]][["AUC"]]
auc.exs[dist]<-exs.sims[[dist]][["AUC"]]
}

auc.is["Method"]="IS"
auc.exs["Method"]="expertsurv"

auc.df2<-bind_rows(auc.is,auc.exs) %>%
pivot_longer(1:5,
names_to="Distribution",
values_to="AUC")

auc.summary2<-auc.df2 %>%
group_by(Distribution,Method) %>%
summarise(mean=mean(AUC),
sd=sd(AUC),
median=median(AUC),
lwr.95=quantile(AUC,0.025),
upr.95=quantile(AUC,0.975))

# Same for AUC
auc.print2 <- auc.summary2 %>%
pivot_wider(names_from = Method,values_from = 3:7) %>%
mutate(mean_diff=mean_IS-mean_expertsurv,
var.ratio=sd_IS^2/sd_expertsurv^2,
Distribution=str_replace_all(Distribution,distributions)) %>%
mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)))  %>%
mutate(expertsurv=paste0(mean_expertsurv," (",lwr.95_expertsurv,", ",upr.95_expertsurv,")"),
IS=paste0(mean_IS," (",lwr.95_IS,", ",upr.95_IS,")"),
Var.Ratio=var.ratio
) %>%
ungroup() %>%
select(Distribution,expertsurv,IS,Mean.Diff=mean_diff,Var.Ratio)

kbl(auc.print2,
caption = "Comparison of probabilistic total area under the curve (mean lifetime OS) estimates, obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, and with the fully Bayesian method of Cooney and White implemented in the expertsurv package. Values shown are probabilstic mean and 95 percent confidence/credible intervals.",
booktabs = T
) %>%
  kable_styling(latex_options=c("striped","HOLD_position"))

# AUC density plot
# No truncation of AUC values but truncation of graph
library(ggridges)
cat('\n\n')
g.auc<-ggplot(data=auc.df2,aes(x=AUC,y=Distribution,colour=Method,fill=Method))+
geom_density_ridges(alpha=0.5,scale=0.95)+
scale_fill_manual(values=cbPalette[2:3])+                                                      scale_colour_manual(values=cbPalette[2:3])+
labs(x="AUC (Mean lifetime OS, Months)",y="Density")+
xlim(0,360)+
theme(legend.position = "bottom")
g.auc
cat('\n\n')
```




## Comparisons of 5-year OS
Table \@ref(tab:ststar) and Figure \@ref(fig:ststar) compare probabilistic estimates of 5-year OS obtained from the two methods.

```{r ststar,  results='asis', fig.cap = "Density plots comparing 5-year OS estimates, obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, with those obtained from the fully Bayesian method of Cooney and White implemented in the expertsurv package."}
library(scales)

surv.tstar2<-data.frame()

for (dist in dists[-6])
{
                                                    tmp.df<-data.frame("expertsurv"=unname(exs.sims[[dist]][["s.tstar"]]), "IS"=unname(is.sims[[dist]][["s.tstar"]]), "dist"=dist)

surv.tstar2<-bind_rows(surv.tstar2,tmp.df)

rm(tmp.df)

}

surv.tstar.long2<-pivot_longer(surv.tstar2,cols=1:2,values_to = "S.tstar") %>% mutate(Output=factor(name,levels=c("IS","expertsurv")), Distribution=str_replace_all(dist,distributions))

surv.tstar.summary2<-surv.tstar.long2 %>%  group_by(Distribution,Output) %>% summarise(mean=mean(S.tstar), median=median(S.tstar), sd=sd(S.tstar), lwr.95=quantile(S.tstar,0.025), upr.95=quantile(S.tstar,0.975)) %>% ungroup %>% mutate(across(where(is.numeric),~percent(.x,accuracy=0.01)))
cat('\n\n')

kbl(surv.tstar.summary2, caption = "Comparison of 5-year OS estimates, obtained from combining trial data with external information using the importance sampling (IS) method described in the paper, with the fully Bayesian method of Cooney and White implemented in the expertsurv package.",
    booktabs=T) %>%
  kable_styling(latex_options=c("striped","HOLD_position"))

g.ststar2<-ggplot(data=surv.tstar.long2,aes(x=S.tstar,y=Distribution,colour=Output,fill=Output))+ geom_density_ridges(alpha=0.5,scale=1)+scale_fill_manual(values=cbPalette)+scale_colour_manual(values=cbPalette)+labs(x="5-year landmark OS",y="Density")
cat('\n\n')
g.ststar2
cat('\n\n')
```





# Scenario analysis: alternative specifications of external information

## Description of included scenarios

The base case analysis assumes a prior distribution for 5-year OS given by normal prior with central 95% density contained within the range (20% to 55%); this distribution has mean ```r mu_t ``` and variance ```r sigma_t^2```. We consider five additional scenarios:

+ Scenario 1: Alternative specification of base case: Beta distribution with same mean and variance 
+ Scenario 2: Base case variance halved ("more confident"), mean unchanged 
+ Scenario 3: Base case variance doubled ("less confident"), mean unchanged
+ Scenario 4: Base case mean decreased by 20%, variance unchanged, Beta distribution ("wrong with same confidence").
+ Scenario 5: Base case mean decreased by 20%, variance halved, Beta distribution ("wrong and more confident")
+ Scenario 6: Base case mean decreased by 20%, variance doubled, Beta distribution ("wrong and less confident")



```{r}

# Prior means for each scenario
mu.ts<-c(mu_t, #BC
         mu_t, #SCA 1...
         mu_t,
         mu_t,
         mu_t-0.2,
         mu_t-0.2,
         mu_t-0.2 #...SCA 6
         )

# Prior variances
var.t<-sigma_t^2
var.ts<-c(1,1,0.5,2,1,0.5,2)*var.t

sigma.ts<-sqrt(var.ts)


# Get params of Beta distributions with given mean and var
estBetaParams <- function(mu, var) {
  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
  beta <- alpha * (1 / mu - 1)
  return(params = list(alpha = alpha, beta = beta))
}

# Scenarios 1, 4, 5 and 6 use beta priors 
beta.params<-estBetaParams(mu.ts[c(2,5:7)],var.ts[c(2,5:7)])

# Distrbution summaries



prior.sca<-data.frame("Distribution"=c("Normal","Beta",rep("Normal",2),rep("Beta",3)),
                                     "mean"=mu.ts,
                                      "sd"=sigma.ts,
                      "Beta.a"=estBetaParams(mu.ts,var.ts)[1],
                      "Beta.b"=estBetaParams(mu.ts,var.ts)[2]
                      ) %>%
  mutate(Beta.a=if_else(Distribution=="Beta",alpha,NA),
         Beta.b=if_else(Distribution=="Beta",beta,NA),
         lwr.95=if_else(Distribution=="Beta",
                        qbeta(0.025,Beta.a,Beta.b),
                        qnorm(0.025,mean=mean,sd=sd)),
         upr.95=if_else(Distribution=="Beta",
                        qbeta(0.975,Beta.a,Beta.b),
                        qnorm(0.975,mean=mean,sd=sd))
                     )


prior.sca.print<-prior.sca %>%
  mutate(Scenario=c("Base case", paste0("Scenario ",1:6)),
         Mean=mean,
         Variance=sd^2,
         across(where(is.numeric),~format(.x,digits=3, nsmall=3)),
         Distribution=if_else(Distribution=="Beta",
                              paste0("Beta(",Beta.a,",",Beta.b,")"),
                              paste0("Normal(",mean,",",sd,"^2)")
                              ),
         Prior.95CrI=paste0("(",lwr.95," to ", upr.95,")"),
         Description = c("NA",
                         "Alternative specification of base case\n using Beta distribution",
                         "More confident",
                         "Less confident",
                         "Wrong with same confidence",
                         "Wrong and more confident",
                         "Wrong and less confident")
         ) %>%
  select(Scenario,Description,Distribution, Mean, Variance,Prior.95CrI)

kbl(prior.sca.print,
    digits=3,
    caption="Summary of external information used in the base case and scenario analysis",
    booktabs=T
    )%>%
  kable_styling(latex_options=c("striped","HOLD_position")) %>%
  column_spec(2, width = "15em")




```


## Results of Scenario analyses:

For all scenarios, we plot posterior mean survival curve estimates, and compare mean lifetime OS and 5-year landmark OS.

```{r}


if(file.exists("is.scenarios.sims.RDS")&file.exists("is.scenarios.RDS"))
   {
     is.scenarios<-readRDS("is.scenarios.RDS")
     is.scenarios.sims<-readRDS("is.scenarios.sims.RDS")
} else{

is.scenarios<-list()
is.scenarios.sims<-list()

for(i in 1:7){
  
  prior.dist<-prior.sca$Distribution[i]
  
  prior.loss<-if(prior.dist=="Normal"){
    function(x){dnorm(x,
                      mean=prior.sca$mean[i],
                      sd=prior.sca$sd[i],
                      log=TRUE)}
  } else{
    function(x){dbeta(x,
                      shape1=prior.sca$Beta.a[i],
                      shape2=prior.sca$Beta.b[i],
                      log=TRUE)
    
  }}
  
  ex.info.sca<-list("tstar"=tstar,
                  "loss"=prior.loss,
                  "lower.probs"=prior.sca$lwr.95[i],
                  "upper.probs"=prior.sca$upr.95[i]
)
  
  is.scenarios[[i]]<-list()
  is.scenarios.sims[[i]]<-list()
  
  for(dist in dists){
  
  old.params<-mle.sims[[dist]]
   # Model fits for later
  fit<-is_surv(
    coeff=old.params$coeff,
    cov=old.params$cov,
    dist=dist,
    ex_info=ex.info.sca
  )
  
  is.scenarios[[i]][[dist]]<-fit
  is.scenarios.sims[[i]][[dist]]<-get_sims(dist=dist,coeff=fit$post_mean,cov=fit$post_cov,tmax=t.max)
  rm(old.params,fit)
  
}
}
saveRDS(is.scenarios,file="is.scenarios.RDS")
saveRDS(is.scenarios.sims,file="is.scenarios.sims.RDS")
}
```




```{r}
# Data frames of simulations

#Number of scenarios incl. base case 
n.sca<-nrow(prior.sca)

#AUC
sims.auc<-data.frame(matrix(NA,nrow=5000,ncol=n.sca*6))
names(sims.auc)<-paste0(c(rep("BC_",6),paste0(rep("SCA",(n.sca-1)*6),rep(1:(n.sca-1),each=6),"_")),
                          rep(dists,n.sca))
# 5 year OS
sims.ststar<-sims.auc

# Extract the simulations
for(i in 1:n.sca){
for(j in seq_along(dists))
{
sims.auc[,(i-1)*6+j]<-is.scenarios.sims[[i]][[dists[j]]][["AUC"]]
sims.ststar[,(i-1)*6+j]<-is.scenarios.sims[[i]][[dists[j]]][["s.tstar"]]
}
}

names(sims.auc)<-paste0(names(sims.auc),"_AUC")
names(sims.ststar)<-paste0(names(sims.ststar),"_STSTAR")

sca.sims<-cbind(sims.auc,sims.ststar)


sca.longsims<-sca.sims%>%
  pivot_longer(1:84,
               names_to="name",
               values_to="value")%>%
  separate_wider_delim(name,delim="_",names=c("Scenario","Distribution","Variable"))

# For later use with ggplot - full distribution and scenario names
sca.longsims2<-sca.longsims%>%
    mutate(Distribution=str_replace_all(Distribution,distributions),
         Scenario=str_replace_all(Scenario,"SCA","Scenario "),
         Scenario=str_replace_all(Scenario,"BC","Base case "))

# Create summary stats
sca.summary<-sca.longsims%>%
  group_by(Scenario,Distribution,Variable) %>%
  summarise(mean=mean(value),
            sd=sd(value),
            median=median(value),
            lwr.95=quantile(value,0.025),
            upr.95=quantile(value,0.975))%>%
  ungroup()%>%
  mutate(across(where(is.character),str_trim))

# Calculate differences versus BC

sca.diffs<-sca.summary %>%
  #mutate(Scenario=str_replace_all(Scenario," ",""))%>%
  select(Distribution, Scenario, Variable, mean, sd) %>%
  pivot_wider(names_from=Scenario,
              values_from=4:5)%>%
  mutate(across(mean_SCA1:mean_SCA6,~.x-mean_BC,.names = "mean.diff_{.col}"),
         across(sd_SCA1:sd_SCA6,~(.x/sd_BC)^2,.names="var.ratio_{.col}")
         ) %>%
  select(Distribution,Variable,contains(c("diff","ratio")))%>%
  rename_with(~gsub("_sd|_mean","",.x)) %>%
  pivot_longer(3:14,names_to = "name",values_to = "value") %>%
  separate_wider_delim(name,delim="_",names=c("Stat","Scenario"))%>%
 # mutate(Scenario=gsub("o","o ",Scenario),
 #        Distribution=str_trim(Distribution)) %>%
  pivot_wider(names_from = Stat,values_from = value)

# Merge with summary table

sca.summary<-sca.summary %>%
    full_join(sca.diffs,by=join_by(Distribution,Scenario,Variable)) %>%
  mutate(Distribution=str_replace_all(Distribution,distributions),
         Scenario=str_replace_all(Scenario,"SCA","Scenario "),
         Scenario=str_replace_all(Scenario,"BC","Base case ")
         )
  

```



```{r,fig.dim=c(7,9), fig.cap="Survival curves (posterior median) for the base case and scenarios 1 through 6",dpi=300}

# Get posterior median survival time estimates
sca.ests<-list()
for(i in 1:n.sca){
  sca.ests[[i]]<-data.frame(matrix(NA,nrow=length(tseq2),ncol=7))
  names(sca.ests[[i]])<-c("time",dists)
  sca.ests[[i]]$time<-tseq2
  
for(j in seq_along(dists))
{
  
  sca.ests[[i]][,dists[j]]<-is.scenarios.sims[[i]][[dists[j]]][["survsummary"]]$S_median
}
}
names(sca.ests)<-c("BC",paste0("SCA_",1:6))

# Reshape for plotting
#
sc.names<-c("Base Case",
            paste0("Scenario ",1:6)
            )
names(sc.names)<-c("BC",
                   paste0("SCA_",1:6))

sca.ests.all<-bind_rows(sca.ests,.id="id") %>%
  pivot_longer(3:8,
               names_to="dist",
               values_to="St") %>%
  mutate(Scenario=str_replace_all(id,sc.names),
         Distribution=str_replace_all(dist,distributions))
  

ggplot(data=sca.ests.all,aes(x=time,y=St,colour=Scenario))+
  geom_line(lwd=1)+
  facet_wrap(.~Distribution,ncol=2)+
  theme(legend.position = "bottom")+
  ylab("S(t)")


```






```{r}

cat('\n\n')

auc.summary<-sca.summary %>%
  filter(Variable=="AUC") %>%
  arrange(Distribution) %>%
  mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)),
         CrI.95 = paste0("(",lwr.95,", ",upr.95,")")
         ) %>%
  select(Distribution,Scenario,Mean=mean,Median=median,CrI.95,Mean.Difference=mean.diff,Variance.Ratio=var.ratio)

kbl(auc.summary,
    caption="Mean OS for Scenarios 1 through 5, including comparisons with base case (probabilistic mean difference and variance ratio)",
    booktabs=T
    )%>%
  kable_styling(latex_options="striped") %>%
  add_header_above(c(" "=5,"Scenario versus Base Case"=2))

#knitr::kable(auc.summary,caption="Mean OS for Scenarios 1 through 5, including comparisons with base case (probabilistic mean difference and variance ratio)")

cat('\n\n')

```

```{r, fig.dim=c(7,9), fig.cap="Boxplots of probabilistic simulations of mean OS in the Base Case and in Scenarios 1 through 5.",dpi=300}

# IN USE


g.auc<-ggplot(data=filter(sca.longsims2,Variable=="AUC"),aes(y=value,x=Scenario,fill=Scenario))+
  geom_boxplot(outliers=FALSE)+
  facet_wrap(.~Distribution,ncol=2,scales="free_y")+
  labs(y="AUC (Mean lifetime OS, Months)")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank())
  
g.auc

cat('\n\n')

```


```{r}

s5.summary<-sca.summary %>%
  filter(Variable=="STSTAR") %>%
  arrange(Distribution) %>%
  mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)),
         CrI.95 = paste0("(",lwr.95,", ",upr.95,")")
         ) %>%
  select(Distribution,Scenario,Mean=mean,Median=median,CrI.95,Mean.Difference=mean.diff,Variance.Ratio=var.ratio)

kbl(s5.summary,
    caption="Summary of probabilistic estimates of 5-year OS for the base case and Scenarios 1 through 5, including comparisons between each scenario and the base case (probabilistic mean difference and variance ratio)",
    booktabs=T
    )%>%
  kable_styling(latex_options="striped")%>%
  add_header_above(c(" "=5,"Scenario versus Base Case"=2))

cat('\n\n')

```

```{r, fig.dim=c(7,9),fig.cap="Boxplots of probabilistic simulations of 5-year OS in the Base Case and in Scenarios 1 through 5.",dpi=150}

# IN USE

g.s5<-ggplot(data=filter(sca.longsims2,Variable=="STSTAR"),aes(y=value,x=Scenario,fill=Scenario))+
  geom_boxplot(outliers=FALSE)+
  facet_wrap(.~Distribution,ncol=2,scales="free_y")+
  labs(y="5-year survival probabilities")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank())
  
g.s5


cat('\n\n')
```


# Assessment of multivariate normal approximation of joint parameter distributions

```{r}

### Compute 5-year OS and mean OS using weighted samples from IS method

wt.summary<-data.frame()

for(dist in dists)
{
  # Extract parameter sims and weights
  raw.samples<-is.models[[dist]]$raw_samples_nat
  raw.samples.mvn<-is.models[[dist]]$raw_samples_mvn
  sample.weights<-is.models[[dist]]$sample_weights
  
  # Predict survival
  # From raw samples
  st.raw<- data.frame("S.tstar"=s_fun(t=tstar,dist=dist,trans_pars = raw.samples),
                      "S.mean"=get_auc(params=raw.samples.mvn,dist=dist,lwr=0,upr = t.max),
                      "Method"="Samples",
                      "Weight"=sample.weights,
                      "Distribution"=dist)
  
  # From summary MVN distribution
  st.mvn<-data.frame("S.tstar"=unname(is.sims[[dist]][["s.tstar"]]),
                     "S.mean"=is.sims[[dist]][["AUC"]],
                     "Method"="MVN",
                     "Weight"=1,
                     "Distribution"=dist)
 
   # Combine into existing datafram
  
  wt.summary<-rbind.data.frame(wt.summary,st.raw,st.mvn)
  
}


##Summarise
wt.summary.long<-wt.summary %>% 
  pivot_longer(1:2) %>% 
  mutate(Distribution=str_replace_all(Distribution,distributions),
         name=str_replace_all(name,c("S.mean"="Mean OS (months)","S.tstar"="5-year OS (prob.)")))

wt.summary.long%>%
  group_by(Distribution,name,Method) %>%
  summarise(mean=weighted.mean(value,w=Weight),
            median=wtd.quantile(value,weights=Weight,probs=0.5),
            lwr.95=wtd.quantile(value,weights=Weight,probs=0.025),
            upr.95=wtd.quantile(value,weights=Weight,probs=0.975)
            ) %>%
  ungroup() %>%
  mutate(across(where(is.numeric),~format(.x,digits=2,nsmall=2)),
         Post.95CrI = paste0("(",lwr.95," to ",upr.95,")")
         ) %>%
  select(-lwr.95,-upr.95) %>%
  pivot_wider(values_from=mean:Post.95CrI,names_from = Method,names_sep=".") %>%
  mutate()%>%
  select(Distribution,Output=name,mean.MVN,median.MVN,Post.95CrI.MVN,mean.Samples,median.Samples,Post.95CrI.Samples) %>%
  kbl(booktabs=T,
      col.names=c("Distribution","Output","Mean","Median","95% CrI","Mean","Median","95% CrI"),
      caption = "Comparison of probabilistic estimates Mean OS (months) and 5-year OS probabilities using the multivariate normal approximation and the weighted samples obtained from the importance sampling algorithm") %>%
  kable_styling(latex_options=c("striped","HOLD_position")) %>%
  add_header_above(c(" "=2,
                     "MVN approximation"=3,
                     "Weighted Samples"=3
                     ))
  
  
```



```{r, fig.cap="Boxplots of probabilistic simulations of Mean OS (months) and 5-year OS probabilities using the multivariate normal approximation and the weighted samples obtained from the importance sampling algorithm.",dpi=300}
  
## Boxplot

ggplot(data=wt.summary.long,aes(y=value,x=Distribution,fill = Method,weight=Weight))+
  geom_boxplot(outliers=FALSE)+
  facet_wrap(.~name,ncol=2,scales="free_y")+
    theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank(),
        axis.title.y=element_blank())


```