---
title: "Supplementary Appendix 4 - additional scenarios"
geometry: margin=2cm
output: 
  bookdown::pdf_document2:
      extra_dependencies: ["float","longtable","booktabs"]
      number_sections: yes
      fig_caption: true
      
---

```{r setup, include=FALSE}

library(tidyverse)
library(flexsurv)
library(GGally)
library(ggpubr)
library(kableExtra)
library(formatR)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE,
                      message= FALSE, 
                      warning=FALSE, 
                      cache=TRUE,
                      tidy=TRUE,
                      dev = 'png',
                      tidy.opts=list(width.cutoff=60),
                      fig.pos = "H", 
                      out.extra = "",
                      latex_options = "HOLD_position",
                      results='asis'
                      )

set.seed(12345)

```


```{r run}
# Fit the models etc
source("setup.R")
source("run_all.R")

```



# Scenarios







The base case analysis assumes a prior distribution for 5-year OS given by normal prior with central 95% density contained within the range (20% to 55%); this distribution has mean ```r mu_t ``` and variance ```r sigma_t^2```. We consider five additional scenarios:

+ Scenario 1: Alternative specification of base case: Beta distribution with same mean and variance 
+ Scenario 2: Base case variance halved ("right and more confident"), mean unchanged yielding 95% prior interval of 25% to 50%
+ Scenario 3: Base case variance doubled ("right and less confident"), mean unchanged 13% to 62%
+ Scenario 4: Base case mean decreased by 20%, variance unchanged, Beta distribution ("wrong with same confidence").
+ Scenario 5: Base case mean decreased by 20%, variance halved, Beta distribution ("wrong and more confident")
+ Scenario 6: Base case mean decreased by 20%, variance doubled, Beta distribution ("wrong and less confident")

For all scenarios, we compare mean lifetime OS and 5-year landmark OS, as before.

```{r}

# Prior means for each scenario
mu.ts<-c(mu_t, #BC
         mu_t, #SCA 1...
         mu_t,
         mu_t,
         mu_t-0.2,
         mu_t-0.2,
         mu_t-0.2 #...SCA 6
         )

# Prior variances
var.t<-sigma_t^2
var.ts<-c(1,1,0.5,2,1,0.5,2)*var.t

sigma.ts<-sqrt(var.ts)


# Get params of Beta distributions with given mean and var
estBetaParams <- function(mu, var) {
  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
  beta <- alpha * (1 / mu - 1)
  return(params = list(alpha = alpha, beta = beta))
}

# Scenarios 1, 4, 5 and 6 use beta priors 
beta.params<-estBetaParams(mu.ts[c(2,5:7)],var.ts[c(2,5:7)])

# Distrbution summaries



prior.sca<-data.frame("Distribution"=c("Normal","Beta",rep("Normal",2),rep("Beta",3)),
                                     "mean"=mu.ts,
                                      "sd"=sigma.ts,
                      "Beta.a"=estBetaParams(mu.ts,var.ts)[1],
                      "Beta.b"=estBetaParams(mu.ts,var.ts)[2]
                      ) %>%
  mutate(Beta.a=if_else(Distribution=="Beta",alpha,NA),
         Beta.b=if_else(Distribution=="Beta",beta,NA),
         lwr.95=if_else(Distribution=="Beta",
                        qbeta(0.025,Beta.a,Beta.b),
                        qnorm(0.025,mean=mean,sd=sd)),
         upr.95=if_else(Distribution=="Beta",
                        qbeta(0.975,Beta.a,Beta.b),
                        qnorm(0.975,mean=mean,sd=sd))
                     )

kbl(prior.sca,digits=3)




```




```{r}


if(file.exists("is.scenarios.sims.RDS")&file.exists("is.scenarios.RDS"))
   {
     is.scenarios<-readRDS("is.scenarios.RDS")
     is.scenarios.sims<-readRDS("is.scenarios.sims.RDS")
} else{

is.scenarios<-list()
is.scenarios.sims<-list()

for(i in 1:7){
  
  prior.dist<-prior.sca$Distribution[i]
  
  prior.loss<-if(prior.dist=="Normal"){
    function(x){dnorm(x,
                      mean=prior.sca$mean[i],
                      sd=prior.sca$sd[i],
                      log=TRUE)}
  } else{
    function(x){dbeta(x,
                      shape1=prior.sca$Beta.a[i],
                      shape2=prior.sca$Beta.b[i],
                      log=TRUE)
    
  }}
  
  ex.info.sca<-list("tstar"=tstar,
                  "loss"=prior.loss,
                  "lower.probs"=prior.sca$lwr.95[i],
                  "upper.probs"=prior.sca$upr.95[i]
)
  
  is.scenarios[[i]]<-list()
  is.scenarios.sims[[i]]<-list()
  
  for(dist in dists){
  
  old.params<-mle.sims[[dist]]
   # Model fits for later
  fit<-is_surv(
    coeff=old.params$coeff,
    cov=old.params$cov,
    dist=dist,
    ex_info=ex.info.sca
  )
  
  is.scenarios[[i]][[dist]]<-fit
  is.scenarios.sims[[i]][[dist]]<-get_sims(dist=dist,coeff=fit$post_mean,cov=fit$post_cov,tmax=t.max)
  rm(old.params,fit)
  
}
}
saveRDS(is.scenarios,file="is.scenarios.RDS")
saveRDS(is.scenarios.sims,file="is.scenarios.sims.RDS")
}
```




```{r}
# Data frames of simulations

#Number of scenarios incl. base case 
n.sca<-nrow(prior.sca)

#AUC
sims.auc<-data.frame(matrix(NA,nrow=5000,ncol=n.sca*6))
names(sims.auc)<-paste0(c(rep("BC_",6),paste0(rep("SCA",(n.sca-1)*6),rep(1:(n.sca-1),each=6),"_")),
                          rep(dists,n.sca))
# 5 year OS
sims.ststar<-sims.auc

# Extract the simulations
for(i in 1:n.sca){
for(j in seq_along(dists))
{
sims.auc[,(i-1)*6+j]<-is.scenarios.sims[[i]][[dists[j]]][["AUC"]]
sims.ststar[,(i-1)*6+j]<-is.scenarios.sims[[i]][[dists[j]]][["s.tstar"]]
}
}

names(sims.auc)<-paste0(names(sims.auc),"_AUC")
names(sims.ststar)<-paste0(names(sims.ststar),"_STSTAR")

sca.sims<-cbind(sims.auc,sims.ststar)


sca.longsims<-sca.sims%>%
  pivot_longer(1:84,
               names_to="name",
               values_to="value")%>%
  separate_wider_delim(name,delim="_",names=c("Scenario","Distribution","Variable"))

# For later use with ggplot - full distribution and scenario names
sca.longsims2<-sca.longsims%>%
    mutate(Distribution=str_replace_all(Distribution,distributions),
         Scenario=str_replace_all(Scenario,"SCA","Scenario "),
         Scenario=str_replace_all(Scenario,"BC","Base case "))

# Create summary stats
sca.summary<-sca.longsims%>%
  group_by(Scenario,Distribution,Variable) %>%
  summarise(mean=mean(value),
            sd=sd(value),
            median=median(value),
            lwr.95=quantile(value,0.025),
            upr.95=quantile(value,0.975))%>%
  ungroup()%>%
  mutate(across(where(is.character),str_trim))

# Calculate differences versus BC

sca.diffs<-sca.summary %>%
  #mutate(Scenario=str_replace_all(Scenario," ",""))%>%
  select(Distribution, Scenario, Variable, mean, sd) %>%
  pivot_wider(names_from=Scenario,
              values_from=4:5)%>%
  mutate(across(mean_SCA1:mean_SCA6,~.x-mean_BC,.names = "mean.diff_{.col}"),
         across(sd_SCA1:sd_SCA6,~(.x/sd_BC)^2,.names="var.ratio_{.col}")
         ) %>%
  select(Distribution,Variable,contains(c("diff","ratio")))%>%
  rename_with(~gsub("_sd|_mean","",.x)) %>%
  pivot_longer(3:14,names_to = "name",values_to = "value") %>%
  separate_wider_delim(name,delim="_",names=c("Stat","Scenario"))%>%
 # mutate(Scenario=gsub("o","o ",Scenario),
 #        Distribution=str_trim(Distribution)) %>%
  pivot_wider(names_from = Stat,values_from = value)

# Merge with summary table

sca.summary<-sca.summary %>%
    full_join(sca.diffs,by=join_by(Distribution,Scenario,Variable)) %>%
  mutate(Distribution=str_replace_all(Distribution,distributions),
         Scenario=str_replace_all(Scenario,"SCA","Scenario "),
         Scenario=str_replace_all(Scenario,"BC","Base case ")
         )
  

```

# Results - survival curves

```{r,fig.dim=c(7,9), fig.cap="Survival curves (posterior median) for the base case and scenarios 1 through 6",dpi=300}

# Get posterior median survival time estimates
sca.ests<-list()
for(i in 1:n.sca){
  sca.ests[[i]]<-data.frame(matrix(NA,nrow=length(tseq2),ncol=7))
  names(sca.ests[[i]])<-c("time",dists)
  sca.ests[[i]]$time<-tseq2
  
for(j in seq_along(dists))
{
  
  sca.ests[[i]][,dists[j]]<-is.scenarios.sims[[i]][[dists[j]]][["survsummary"]]$S_median
}
}
names(sca.ests)<-c("BC",paste0("SCA_",1:6))

# Reshape for plotting
#
sc.names<-c("Base Case",
            paste0("Scenario ",1:6)
            )
names(sc.names)<-c("BC",
                   paste0("SCA_",1:6))

sca.ests.all<-bind_rows(sca.ests,.id="id") %>%
  pivot_longer(3:8,
               names_to="dist",
               values_to="St") %>%
  mutate(Scenario=str_replace_all(id,sc.names),
         Distribution=str_replace_all(dist,distributions))
  

ggplot(data=sca.ests.all,aes(x=time,y=St,colour=Scenario))+
  geom_line(lwd=1)+
  facet_wrap(.~Distribution,ncol=2)+
  theme(legend.position = "bottom")+
  ylab("S(t)")


```

# Results - mean lifetime OS




```{r}

cat('\n\n')

auc.summary<-sca.summary %>%
  filter(Variable=="AUC") %>%
  arrange(Distribution) %>%
  mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)),
         CrI.95 = paste0("(",lwr.95,", ",upr.95,")")
         ) %>%
  select(Distribution,Scenario,Mean=mean,Median=median,CrI.95,Mean.Difference=mean.diff,Variance.Ratio=var.ratio)

kbl(auc.summary,
    caption="Mean OS for Scenarios 1 through 5, including comparisons with base case (probabilistic mean difference and variance ratio)",
    booktabs=T
    )%>%
  kable_styling(latex_options="striped") %>%
  add_header_above(c(" "=5,"Scenario versus Base Case"=2))

#knitr::kable(auc.summary,caption="Mean OS for Scenarios 1 through 5, including comparisons with base case (probabilistic mean difference and variance ratio)")

cat('\n\n')

```

```{r, fig.dim=c(7,9), fig.cap="Boxplots of probabilistic simulations of mean OS in the Base Case and in Scenarios 1 through 5.",dpi=300}

# IN USE


g.auc<-ggplot(data=filter(sca.longsims2,Variable=="AUC"),aes(y=value,x=Scenario,fill=Scenario))+
  geom_boxplot(outliers=FALSE)+
  facet_wrap(.~Distribution,ncol=2,scales="free_y")+
  labs(y="AUC (Mean lifetime OS, Months)")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank())
  
g.auc

cat('\n\n')

```
# Results - 5 year survival probabilities

```{r}

s5.summary<-sca.summary %>%
  filter(Variable=="STSTAR") %>%
  arrange(Distribution) %>%
  mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)),
         CrI.95 = paste0("(",lwr.95,", ",upr.95,")")
         ) %>%
  select(Distribution,Scenario,Mean=mean,Median=median,CrI.95,Mean.Difference=mean.diff,Variance.Ratio=var.ratio)

kbl(s5.summary,
    caption="Summary of probabilistic estimates of 5-year OS for the base case and Scenarios 1 through 5, including comparisons between each scenario and the base case (probabilistic mean difference and variance ratio)",
    booktabs=T
    )%>%
  kable_styling(latex_options="striped")%>%
  add_header_above(c(" "=5,"Scenario versus Base Case"=2))

cat('\n\n')

```

```{r, fig.dim=c(7,9),fig.cap="Boxplots of probabilistic simulations of 5-year OS in the Base Case and in Scenarios 1 through 5.",dpi=150}

# IN USE

g.s5<-ggplot(data=filter(sca.longsims2,Variable=="STSTAR"),aes(y=value,x=Scenario,fill=Scenario))+
  geom_boxplot(outliers=FALSE)+
  facet_wrap(.~Distribution,ncol=2,scales="free_y")+
  labs(y="5-year survival probabilities")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank())
  
g.s5


cat('\n\n')
```

```{r, cache=FALSE}

knitr::knit_exit()
```


# Old code - not in use

# Results - mean lifetime OS



```{r}

#AUC for the models
auc.is.sca<-data.frame(matrix(NA,nrow=5000,ncol=6*6))

names(auc.is.sca)<-paste0(c(rep("Base case : ",6),paste0(rep("Scenario ",30),rep(1:5,each=6)," : ")),
                          rep(distributions,6))

for(i in 1:6){
for(j in seq_along(dists))
{
auc.is.sca[,(i-1)*6+j]<-is.scenarios.sims[[i]][[dists[j]]][["AUC"]]

}
}


auc.sca.long<-auc.is.sca %>%
pivot_longer(1:36,
names_to="name",
values_to="AUC")%>%
  separate_wider_delim(name,delim=":",names=c("Scenario","Distribution")) 

# Summarise by distribution and scenario
auc.sca.summary<-auc.sca.long%>%
  group_by(Scenario,Distribution) %>%
  summarise(mean=mean(AUC),
            sd=sd(AUC),
            median=median(AUC),
            lwr.95=quantile(AUC,0.025),
            upr.95=quantile(AUC,0.975))%>%
  ungroup()%>%
  mutate(across(where(is.character),str_trim))

#print(knitr::kable((auc.sca.summary)))




# Calculate difference from base case (mean difference and variance ratio)
# Not used for now

auc.sca.diff<-auc.sca.summary %>%
  mutate(Scenario=str_replace_all(Scenario," ",""))%>%
  select(Distribution, Scenario, mean, sd) %>%
  pivot_wider(names_from=Scenario,
              values_from=3:4)%>%
  mutate(across(3:7,~.x-mean_Basecase,.names = "diff_{.col}"),
         across(9:13,~(.x/sd_Basecase)^2,.names="var_ratio_{.col}"),
         across(where(is.numeric),~format(round(.x,2),nsmall=2))
         ) %>%
  select(Distribution,contains(c("diff","ratio")))%>%
  rename_with(~gsub("_sd","",.x))
  
#print(knitr::kable((auc.sca.diff)))


# In use:
# Second attempt at formatting the differences table
auc.sca.diff2<-auc.sca.summary %>%
  mutate(Scenario=str_replace_all(Scenario," ",""))%>%
  select(Distribution, Scenario, mean, sd) %>%
  pivot_wider(names_from=Scenario,
              values_from=3:4)%>%
  mutate(across(3:7,~.x-mean_Basecase,.names = "mean.diff_{.col}"),
         across(9:13,~(.x/sd_Basecase)^2,.names="var.ratio_{.col}")
         ) %>%
  select(Distribution,contains(c("diff","ratio")))%>%
  rename_with(~gsub("_sd|_mean","",.x)) %>%
  pivot_longer(2:11,names_to = "name",values_to = "value") %>%
  separate_wider_delim(name,delim="_",names=c("Variable","Scenario"))%>%
  mutate(Scenario=gsub("o","o ",Scenario),
         Distribution=str_trim(Distribution)) %>%
  pivot_wider(names_from = Variable,values_from = value)


# Merge with summary and print nicely

auc.sca.summary2<-auc.sca.summary %>%
    full_join(auc.sca.diff2,by=join_by(Distribution,Scenario)) %>%
  arrange(Distribution) %>%
  mutate(across(where(is.numeric),~format(round(.x,2),nsmall=2)),
         CrI.95 = paste0("(",lwr.95,", ",upr.95,")")
         ) %>%
  select(Distribution,Scenario,mean,median,CrI.95,mean.diff,var.ratio)

print(knitr::kable(auc.sca.summary2))

# Plots

# AUC density plot

# NOT USED
# No truncation of AUC values but truncation of graph
library(ggridges)
cat('\n\n')
g.auc<-ggplot(data=auc.sca.long,aes(x=AUC,y=Distribution,colour=Scenario,fill=Scenario))+
geom_density_ridges(alpha=0.5,scale=0.95)+
#scale_fill_manual(values=cbPalette[2:3])+                                                      
#scale_colour_manual(values=cbPalette[2:3])+
labs(x="AUC (Mean lifetime OS, Months)",y="Density")+
xlim(0,180)+
theme(legend.position = "bottom")
  
#g.auc
cat('\n\n')
```

```{r, fig.dim=c(7,9)}

# IN USE

g.auc3<-ggplot(data=auc.sca.long,aes(y=AUC,x=Scenario,fill=Scenario))+
geom_boxplot(outliers=FALSE)+
#scale_fill_manual(values=cbPalette[2:3])+                                                      
#scale_colour_manual(values=cbPalette[2:3])+
facet_wrap(.~Distribution,ncol=2,scales="free_y")+
labs(y="AUC (Mean lifetime OS, Months)")+
theme(legend.position = "none",
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  

g.auc3



```

# Results - 5 year survival probabilities

```{r}

#AUC for the models
s5.is.sca<-data.frame(matrix(NA,nrow=5000,ncol=6*6))

names(s5.is.sca)<-paste0(c(rep("Base case : ",6),paste0(rep("Scenario ",30),rep(1:5,each=6)," : ")),
                          rep(distributions,6))

for(i in 1:6){
for(j in seq_along(dists))
{
s5.is.sca[,(i-1)*6+j]<-is.scenarios.sims[[i]][[dists[j]]][["s.tstar"]]

}
}


s5.sca.long<-s5.is.sca %>%
pivot_longer(1:36,
names_to="name",
values_to="S.tstar")%>%
  separate_wider_delim(name,delim=":",names=c("Scenario","Distribution")) 

# Summarise by distribution and scenario
s5.sca.summary<-s5.sca.long%>%
  group_by(Scenario,Distribution) %>%
  summarise(mean=mean(S.tstar),
            sd=sd(S.tstar),
            median=median(S.tstar),
            lwr.95=quantile(S.tstar,0.025),
            upr.95=quantile(S.tstar,0.975))%>%
  ungroup()

print(knitr::kable((s5.sca.summary)))

# Calculate difference from base case (mean difference and variance ratio)
s5.sca.diff<-s5.sca.summary %>%
  mutate(Scenario=str_replace_all(Scenario," ",""))%>%
  select(Distribution, Scenario, mean, sd) %>%
  pivot_wider(names_from=Scenario,
              values_from=3:4)%>%
  mutate(across(3:7,~.x-mean_Basecase,.names = "diff_{.col}"),
         across(9:13,~(.x/sd_Basecase)^2,.names="var_ratio_{.col}"),
         across(where(is.numeric),~format(round(.x,4),nsmall=4))
         ) %>%
  select(Distribution,contains(c("diff","ratio")))%>%
  rename_with(~gsub("_sd","",.x))
  
print(knitr::kable((s5.sca.diff)))



# Plots

# AUC density plot
# No truncation of AUC values but truncation of graph
library(ggridges)
cat('\n\n')
g.tstar<-ggplot(data=s5.sca.long,aes(x=S.tstar,y=Distribution,colour=Scenario,fill=Scenario))+
geom_density_ridges(alpha=0.5,scale=0.95)+
#scale_fill_manual(values=cbPalette[2:3])+                                                      
#scale_colour_manual(values=cbPalette[2:3])+
labs(x="5-year OS probability",y="Density")+
xlim(0,1)+
theme(legend.position = "bottom")
  
g.tstar
cat('\n\n')






```